Place of death in NRS COVID-19 deaths

### 2021-03-03 covid_nrs_place_of_death.Rmd
updated the table of annual deahts & comparing with past rates

@@

```{r}
## this is fairly ad-hoc... but it works!
table_comparing_rates <-
  bind_rows(
    ## all available data
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               str_starts(place_of_death, "Home")) %>% group_by(year, week_number) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% left_join(
        past_deaths_overall %>% filter(str_starts(place_of_death, "Home") &
                                         year != "average") %>% group_by(week_number) %>% summarise(
                                           average_weekly_deaths = sum(number_of_deaths) / 5,
                                           .groups = "drop"
                                         ),
        by = "week_number"
      ) %>% summarise(
        period = "all available data",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        past_annual_deaths = sum(average_weekly_deaths)
      )
    ,
    ## 2020 only
    bind_cols(
      merged_deaths_overall %>%
        filter(
          age == "all" &
            sex == "all" & str_starts(place_of_death, "Home") & year == 2020
        ) %>%
        summarise(
          period = "2020",
          deaths_all_causes = sum(deaths_all_causes),
          deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
          deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE)
        ),
      past_deaths_overall %>% filter(str_starts(place_of_death, "Home") &
                                       year != "average") %>% summarise(past_annual_deaths = sum(number_of_deaths) / 5)
    ),
    ## 2020 from week 12 onwards (including week 12) - pandemic part of 2020
    bind_cols(
      merged_deaths_overall %>%
        filter(
          age == "all" &
            sex == "all" &
            str_starts(place_of_death, "Home") &
            year == 2020 & week_number >= 12
        ) %>%
        summarise(
          period = "2020, from wk12",
          deaths_all_causes = sum(deaths_all_causes),
          deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
          deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE)
        ),
      past_deaths_overall %>% filter(
        week_number >= 12 &
          str_starts(place_of_death, "Home") &
          year != "average"
      ) %>% summarise(past_annual_deaths = sum(number_of_deaths) / 5)
    )
  ) %>%
  mutate(
    ratio_deaths_all_causes_to_historical = deaths_all_causes / past_annual_deaths,
    ratio_non_covid_deaths_to_historical = deaths_non_covid / past_annual_deaths,
    proportion_covid_deaths = deaths_covid_related / deaths_all_causes
  )

write_csv(x = table_comparing_rates, file = file.path(dir_outputs, "annual_death_rates_comparison_with_historical_deaths.csv"))

table_comparing_rates %>% knitr::kable(caption = "Comparing annual home death rate during COVID pandemic to historical deaths at home")
```

@@

## constants for use in graphs & annotations
weeks_available_2021 <- max(merged_deaths_age$week_number[merged_deaths_age$year==2021])
most_recent_date_available_2021 <- max(merged_deaths_age$date_w_c)
earliest_date <- min(merged_deaths_age$date_w_c)

>>>>

# Constants about latest data available

```{r}
## constants for use in graphs & annotations
weeks_available_2021 <- max(merged_deaths_age$week_number[merged_deaths_age$year==2021])
most_recent_date_available_2021 <- max(merged_deaths_age$date_w_c)
earliest_date <- min(merged_deaths_age$date_w_c)
```

@@

```{r}
table_current_deaths_compared_to_past <-
  bind_rows(
    ## all available data
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all"
             ) %>% 
      group_by(year, place_of_death, week_number) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        # deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
      left_join(
        past_deaths_overall %>% filter(year != "average") %>% group_by(week_number, place_of_death) %>% summarise(
                                           average_weekly_deaths = sum(number_of_deaths) / 5,
                                           .groups = "drop"
                                         ),
        by = c("week_number", "place_of_death")
      ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
        .groups = "drop"
      )
    ,
    ## 2020 only
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               year == 2020
             ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = "2020",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_overall %>% 
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean),
      by = "place_of_death"
    )
    ,
    ## 2020 from week 12 onwards (including week 12) - pandemic part of 2020
    merged_deaths_overall %>%
      filter(
        age == "all" &
          sex == "all" &
          year == 2020 & week_number >= 12
      ) %>%
      group_by(place_of_death) %>%
      summarise(
        period = "2020, week 12 - week 53",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      past_deaths_overall %>% 
        filter(week_number >= 12 & year != "average") %>%
        group_by(place_of_death) %>%
        summarise(mean_annual_deaths_2015_2019 = sum(number_of_deaths) / 5),
      by = "place_of_death"
    )
  ) %>%
  mutate(
    ratio_deaths_all_causes_to_historical = deaths_all_causes / mean_annual_deaths_2015_2019,
    ratio_non_covid_deaths_to_historical = deaths_non_covid / mean_annual_deaths_2015_2019,
    proportion_covid_deaths = deaths_covid_related / deaths_all_causes
  )

write_csv(x = table_current_deaths_compared_to_past, file = file.path(dir_outputs, "annual_death_rates_comparison_with_historical_deaths.csv"))

table_current_deaths_compared_to_past %>% knitr::kable(caption = "Comparing annual home death rate during COVID pandemic to historical deaths at home")
```

@@
